### 2015-05-27 Kokkos_PG_Compilation.tex
Updated to the document. This version will now be send to R&A.

@@

@@

The current version of NVCC has some shortcomings when used as the main compiler for a project, in particular when part of a complex build system.
For example, it does not understand most GCC command-line options, which must be prepended by the \lstinline|-Xcompiler| flag when calling NVCC. 
Kokkos comes with a shell script, called \lstinline|nvcc_wrapper|, that wraps NVCC to address these issues.
We intend this as a drop-in replacement for a normal GCC-compatible compiler (e.g., GCC or Intel) in your build system.
It analyzes the provided command-line options and prepends them correctly. 
It also adds the correct flags for compiling generic C++ files containing CUDA code (e.g., \verb!*.cpp!, \verb!*.cxx!, or \verb!*.CC!).
By default \lstinline|nvcc_wrapper| calls \verb!g++! as the host compiler.
You may override this by providing NVCC's '\lstinline|-ccbin|' option as a compiler flag.
The default can be reset by editing the script itself.

@@

\section{Building for Intel Xeon Phi}\label{S:build:Phi}

TODO

>>>>

\hline
   KOKKOS\_CUDA\_OPTIONS (IN): & Enable optional settings specific to CUDA. \\
      \hspace{0.5cm}Options: & force\_uvm,use\_ldg,rdc \\
      \hspace{0.5cm}Default: & \\

@@

Kokkos provides a script \lstinline|generate_makefile.bash| which can generate a Makefile for building and installing the library
as well as building and running the tests. Please run \lstinline|generate_makefile.bash --help| for options. 
Note that paths given to the script must be absolute paths, and the script must be run with the \lstinline|bash| shell (the script will
do it if it is run directly i.e. as \lstinline|./generate_makefile.bash|.

@@

The current version of NVCC has some shortcomings when used as the main compiler for a project, in particular when part of a complex build system.
For example, it does not understand most GCC command-line options, which must be prepended by the \lstinline|-Xcompiler| flag when calling NVCC. 
Kokkos comes with a shell script, called \lstinline|nvcc_wrapper|, that wraps NVCC to address these issues.
We intend this as a drop-in replacement for a normal GCC-compatible compiler (e.g., GCC or Intel) in your build system.
It analyzes the provided command-line options and prepends them correctly. 
It also adds the correct flags for compiling generic C++ files containing CUDA code (e.g., \verb!*.cpp!, \verb!*.cxx!, or \verb!*.CC!).
By default \lstinline|nvcc_wrapper| calls \verb!g++! as the host compiler.
You may override this by providing NVCC's '\lstinline|-ccbin|' option as a compiler flag.
The default can be set by editing the script itself or by setting the environment variable \lstinline|NVCC_WRAPPER_DEFAULT_COMPILER|.

@@

### 2015-05-27 Kokkos_PG_HierarchicalParallelism.tex
Updated style of the Guide . This comments out the SAND Report style
The stuff is still in if we really need to its not that hard to go back.
Updated to the document. This version will now be send to R&A.

Kokkos exposes scratch pads through a special memory space associated with the execution space:
\lstinline|execution_space::scratch_memory_space|.
You may allocate a chunk of scratch memory through the \lstinline|TeamPolicy| member type.
You may request multiple allocations from scratch, up to a user-provided maximum. 
The maximum is provided either through a function in the functor which returns a potentially team-size dependent value, 
or it can be specified as an argument to the constructor of the TeamPolicy. 
It is not valid to provide both values at the same time. 
The argument to the TeamPolicy can be used to set the shared memory size when using functors. 
One restriction on shared memory allocations is that they can not be freed during the lifetime of the team. 
This avoids the complexity of a memory pool,
and reduces the time it takes to obtain an allocation
(which currently is a few operations to calculate the offset).

@@

Instead of simply getting raw allocations in memory, users can also allocate Views directly in scratch memory. 
This is achieved by providing the shared memory handle as the first argument of the View constructor.
Views also have a static member function which return their shared memory size requirements. 
The function expects the run-time dimensions as arguments, corresponding to View's constructor.

\begin{lstlisting}
tyepdef Kokkos::DefaultExecutionSpace::scratch_memory_space
  ScratchSpace;
// Define a view type in ScratchSpace
typedef Kokkos::View<int*[4],ScratchSpace> shared_int_2d;

@@

\subsection{Restrict execution with \lstinline|single|}

>>>>

Kokkos exposes scratch pads through a special memory space associated with the execution space:
\lstinline|execution_space::scratch_memory_space|.
You may allocate a chunk of scratch memory through the \lstinline|TeamPolicy| member type.
You may request multiple allocations from scratch, up to a user-provided maximum. 
The maximum is provided either through a function in the functor which returns a potentially team-size dependent value, 
or it can be specified as an argument to the constructor of the TeamPolicy. 
It is not valid to provide both values at the same time. 
The argument to the TeamPolicy can be used to set the shared memory size when using functors. 
One restriction on shared memory allocations is that they can not be freed during the lifetime of the team. 
This avoids the complexity of a memory pool,
and reduces the time it takes to obtain an allocation
(which currently is a few tens of integer operations to calculate the offset).

@@

Instead of simply getting raw allocations in memory, users can also allocate Views directly in scratch memory. 
This is achieved by providing the shared memory handle as the first argument of the View constructor.
Views also have a static member function which return their shared memory size requirements. 
The function expects the run-time dimensions as arguments, corresponding to View's constructor. 
Note that the view must be unmanaged (i.e. it must have the \lstinline|Unmanaged| memory trait).

\begin{lstlisting}
tyepdef Kokkos::DefaultExecutionSpace::scratch_memory_space
  ScratchSpace;
// Define a view type in ScratchSpace
typedef Kokkos::View<int*[4],ScratchSpace,Kokkos::MemoryTraits<Kokkos::Unmanaged>> shared_int_2d;

@@

\subsection{Restricting execution to a single executor}

### 2015-05-27 Kokkos_PG_LegacyCodeIntegration.tex
Updated to the document. This version will now be send to R&A.

\begin{lstlisting}
template<class Scalar, class Device>
Scalar dot(View<const Scalar* , Device> a, 
           View<const Scalar*, Device> b) {
// Check for Cuda memory and call cublas if true
#ifdef KOKKOS_HAVE_CUDA
  if(Impl::is_same<typename Device::memory_space,
                            CudaSpace>::value ||
     Impl::is_same<typename Device::memory_space,
                            CudaUVMSpace>::value) {
    return call_cublas_dot(a.ptr_on_device(), b.ptr_on_device(),
                           a.dimension_0() );
  }
#endif

// Call CBlas on the host otherwise
  if(Impl::is_same<typename Device::memory_space,HostSpace>::value) {
    return call_cblas_dot(a.ptr_on_device(), b.ptr_on_device(),
                          a.dimension_0() );
  }
}
\end{lstlisting}

Note the \lstinline|Impl::is_same| construct can be replaced

>>>>

\begin{lstlisting}
template<class Scalar, class Device>
Scalar dot(View<const Scalar* , Device> a, 
           View<const Scalar*, Device> b) {
// Check for Cuda memory and call cublas if true
#ifdef KOKKOS_HAVE_CUDA
  if(std::is_same<typename Device::memory_space,
                           CudaSpace>::value ||
     std::is_same<typename Device::memory_space,
                           CudaUVMSpace>::value) {
    return call_cublas_dot(a.ptr_on_device(), b.ptr_on_device(),
                           a.dimension_0() );
  }
#endif

// Call CBlas on the host otherwise
  if(std::is_same<typename Device::memory_space,HostSpace>::value) {
    return call_cblas_dot(a.ptr_on_device(), b.ptr_on_device(),
                          a.dimension_0() );
  }
}
\end{lstlisting}

### 2015-05-27 Kokkos_PG_ParallelDispatch.tex
Updated to the document. This version will now be send to R&A.

@@

Kokkos lets users choose whether to use a functor or a lambda.
Lambdas are convenient for short loop bodies.
For a much more complicated loop body,
you might find it easier for testing to separate it out and name it as a functor.
Lambdas by definition are ``anonymous functions,'' meaning that they have no name.
This makes it harder to test them.
Furthermore, if you would like to use lambdas with CUDA,
you must have a sufficiently new version of CUDA.
Finally, the ``execution tag'' feature, which lets you put together several different parallel loop bodies into a single functor,
only works with functors.  (See Chapter \ref{C:Hierarchical} for details.)

@@

The most common parallel dispatch operation is a \lstinline|parallel_for| call.
It corresponds to the OpenMP construct \lstinline!#pragma omp parallel for!.
Parallel for splits the index range over the available hardware resources and executes the loop body in parallel.
Each iteration is executed independently.
Kokkos promises nothing about the loop order or the amount of work which actually runs concurrently. 
This means in particular that not all loop iterations are active at the same time.  Consequently, it is not legal to use wait constructs (e.g., wait for a prior iteration to finish). 
Kokkos also doesn't guarantee that it will use all available parallelism. 
For example, it can decide to execute in serial if the loop count is very small, and it would typically be faster to run in serial instead of introducing parallelization overhead.

@@

>>>>

It is a violation of Kokkos semantics to capture by reference \lstinline|[&]| for two reasons.
First Kokkos might give the lambda to an execution space which can not access the stack of 
the dispatching thread. Secondly, capturing by reference allows the programmer to violate the 
const semantics of the lambda. For correctness and portability reasons lambdas and functors are
treated as const objects inside the parallel code section. Capturing by reference allows a circumvention
of that const property, and enables many more possibilities of writing non-threads-safe code.

When using lambdas for nested parallelism (see Chapter \ref{C:Hierarchical} for details) using
capture by reference can be useful for performance reasons, but the code is only valid Kokkos
code if it also works with capturing by copy.

@@

Kokkos lets users choose whether to use a functor or a lambda.
Lambdas are convenient for short loop bodies.
For a much more complicated loop body,
you might find it easier for testing to separate it out and name it as a functor.
Lambdas by definition are ``anonymous functions,'' meaning that they have no name.
This makes it harder to test them.
Furthermore, if you would like to use lambdas with CUDA,
you must have a sufficiently new version of CUDA (at the time of writing Lambdas are not supported yet supported by the release 7.0).
Finally, the ``execution tag'' feature, which lets you put together several different parallel loop bodies into a single functor,
only works with functors.  (See Chapter \ref{C:Hierarchical} for details.)

@@

The most common parallel dispatch operation is a \lstinline|parallel_for| call.
It corresponds to the OpenMP construct \lstinline!#pragma omp parallel for!.
Parallel for splits the index range over the available hardware resources and executes the loop body in parallel.
Each iteration is executed independently.
Kokkos promises nothing about the loop order or the amount of work which actually runs concurrently. 
This means in particular that not all loop iterations are active at the same time.  Consequently, it is not legal to use wait constructs (e.g., wait for a prior iteration to finish). 
Kokkos also doesn't guarantee that it will use all available parallelism. 
For example, it can decide to execute in serial if the loop count is very small, and it would typically be faster to run in serial instead of introducing parallelization overhead.
The \lstinline|RangePolicy| allows you to specify minimal chunk sizes in order to control potential concurrency for low trip count loops.

@@

This approach can also be used to template the operators by templating the tag classes which is useful to enable compile time evaluation of appropriate conditionals.

### 2015-05-27 Kokkos_PG_Views.tex
Updated to the document. This version will now be send to R&A.

@@

@@

Deep copies can only be performed between views with an identical 
memory layout and padding. For example the following two operations
are not valid:
\begin{lstlisting}
View<int*[3]> a ("a", CudaSpace, 10);
View<int*[3]> b ("b", HostSpace, 10);
deep_copy (a, b); // This will give a compiler error

View<int*[3], LayoutLeft> c ("c", CudaSpace, 10);
View<int*[3], LayoutLeft> d ("d", HostSpace, 10);
deep_copy (c, d); // This might give a runtime error
\end{lstlisting}

@@

\begin{lstlisting}
View<int*[3]> a ("a", MemorySpace, 10);
// Allocate a view in HostSpace with the 
// layout and padding of a
typename View<int*[3]>::HostMirror b =
    create_mirror(a);
// This is always a memcopy
deep_copy (a, b);

@@

The \lstinline!Atomic! access trait lets you create a View of data , such that every
read or write to any entry uses an atomic update.  This only works for
certain data types, for which the hardware and programming model
support atomic updates.  Currently, Kokkos limits atomic updates to
32-bit or 64-bit types\footnote{On X86 architectures Kokkos also supports 
128-bit atomics. Types for which atomic access are performed must support the 
necessary operators such as =,+=,-=,+,- etc.}.

Note that it is a violation of the semantics to write \emph{or} read from a 
non-atomic view of data if at the same time an atomic view of the data is used.

@@

\begin{lstlisting}
// Compute a sparse matrix-vector product, for a sparse
// matrix stored in compressed sparse row (CSR) format.
void
spmatvec (const View<double*>& y, 
  const View<const size_t*>& ptr,
  const View<const int*>& ind,
  const View<const double*>& val,
  const View<const double*>& x)
{
  // Access to x has less locality than access to y.
  View<const double*, RandomAccess> x_ra;
  typedef View<const size_t*>::size_type size_type;

>>>>

Note that the above used constructor is not necessarily available
for all view types. Specific Layouts or Memory Spaces may require more 
specialized allocators. This will be discussed later.

@@

Note the return type of \lstinline|dimension_N()| is the \lstinline|size_type| of the 
views memory space. This causes some issues if warning free compilation
should be achieved since it will typically be necessary to cast the return value.
In particular in cases where the \lstinline|size_type| is more conservative than 
required it can be beneficial to cast the value to \lstinline|int|, since signed 32 bit
integers typically give the best performance when used as index types. In index heavy 
codes this performance difference can be significant compared to using \lstinline|size_t| 
since the vector length on many architectures is twice as long for 32 bit values as for
64 bit values, and signed integers have less stringent overflow testing requirements than
unsigned integers.

@@

Deep copies can only be performed between views with an identical 
memory layout and padding. For example the following two operations
are not valid:
\begin{lstlisting}
View<int*[3], CudaSpace> a ("a", 10);
View<int*[3], HostSpace> b ("b", 10);
deep_copy (a, b); // This will give a compiler error

View<int*[3], LayoutLeft, CudaSpace> c ("c", 10);
View<int*[3], LayoutLeft, HostSpace> d ("d", 10);
deep_copy (c, d); // This might give a runtime error
\end{lstlisting}

@@

\begin{lstlisting}
View<int*[3], MemorySpace> a ("a", 10);
// Allocate a view in HostSpace with the 
// layout and padding of a
typename View<int*[3], MemorySpace>::HostMirror b =
    create_mirror(a);
// This is always a memcopy
deep_copy (a, b);

@@

The \lstinline!Atomic! access trait lets you create a View of data , such that every
read or write to any entry uses an atomic update. Kokkos supports atomics for 
all data types independent of size. Restrictions are that you are (i) not allowed to alias data
for which atomic operations are performed, and the results of non atomic accesses (
including read) to data which is at the same time atomically accessed is not defined.

Performance characteristics of atomic operation depend on the data type. 
Some types (in particular integer types) are natively supported and might even provide 
asynchronous atomic operations. Others (such as 32 bit and 64 bit atomics for non 
integer types) are often implemented using CAS loops of integers. 
Everything else is implemented with a locking approach, where an atomic operations 
aquires a lock based on a hash of the pointer value of the data element.

Types for which atomic access are performed must support the 
necessary operators such as =,+=,-=,+,- etc. as well as have a number of \lstinline|volatile|
overloads of functions such as assign and copy constructors defined.

@@

\begin{lstlisting}
// Compute a sparse matrix-vector product, for a sparse
// matrix stored in compressed sparse row (CSR) format.
void
spmatvec (const View<double*>& y, 
  const View<const size_t*>& ptr,
  const View<const int*>& ind,
  const View<const double*>& val,
  const View<const double*>& x)
{
  // Access to x has less locality than access to y.
  View<const double*, RandomAccess> x_ra = x;
  typedef View<const size_t*>::size_type size_type;
